Dark Control:\\ Towards a Unified Account of Default Mode Function\\ by Markov Decision Processes

### 2017-12-06 paper/paper.tex
Add MDP def at beginning

Psychologically,
the more the ongoing executed task is unknown and unpracticed,
the less stimulus-independent thoughts occur
\citep{filler1973daydreaming, teasdale1995stimulus, christoff2016mind}.
Conversely,
it has been empirically shown that, the more the world is easy to foresee,
the more human mental activity
becomes detached from the actual sensory environment
\citep{antrobus1966studies, pope1978regulation, mason2007, weiss2006}.
Without requiring explicit awareness,
these ``offline'' processes may contribute
to optimizing control of the organism.
We formalize
a \textit{policy matrix} to capture the space of possible actions that
the agent can perform
on the environment given the current state. A \textit{value function}
maps environmental objects and events (i.e., states) to expected rewards.
Switching between states reduces to a sequential processing model.
Informed by outcomes of performed actions, neural computation reflected in
DMN dynamics could be constantly shaped by prediction error through feedback loops.
Such an RL account of DMN function can naturally embed human behavior
into the tension between exploitative action with immediate gains and
exploratory action with longer-term gratification.

@@

Additionally,
this network was proposed to be situated
at the top of the brain network hierarchy, with
the subordinate salience and dorsal attention network in the middle and
the unimodal sensory cortices at the bottom
\citep{carhart2010default, margulies2016situating}.
Its putative involvement in thinking about the past,
hypothetical experiences, and the future
appears to tie in with the implicit computation of
action and state cascades as a function of what happened in the past.
A policy matrix encapsulates the repertoire of possible actions
on the world given a current state.
The policy matrix encodes the probabilites of
choosing actions to be executed in a certain situation.
The DMN may subserve
  constant exploration of possible action trajectories and
  nested estimaton of their
  cumulative reward outcomes. Implicit computation of future choices
  provides an explanation for the
  evolutionary emergence and practical usefulness of
  mind-wandering at day-time and dreams during sleep
  in humans.
\begin{figure}[!h]
  \includegraphics[width=.9\linewidth]{neurobiological_and_rl_overview_DMN.pdf}
  \caption{\textbf{Default mode network:
  neurobiological implementation of reinforcement learning.}
  Overview of how the constituent nodes of the DMN (refer to section \ref{sec:nodes}) may
  map onto computational components necessary for an RL agent.}
  \label{fig:rl_process_chart}
\end{figure}

>>>>

Psychologically,
the more the ongoing executed task is unknown and unpracticed,
the less stimulus-independent thoughts occur
\citep{filler1973daydreaming, teasdale1995stimulus, christoff2016mind}.
Conversely,
it has been empirically shown that, the more the world is easy to foresee,
the more human mental activity
becomes detached from the actual sensory environment
\citep{antrobus1966studies, pope1978regulation, mason2007, weiss2006}.
Without requiring explicit awareness,
these ``offline'' processes may contribute
to optimizing control of the organism.
We formalize
a \textit{policy matrix} to capture the space of possible actions that
the agent can perform
on the environment given the current state. A \textit{value function}
maps environmental objects and events (i.e., states) to expected rewards.
Switching between states reduces to a sequential processing model.
Informed by outcomes of performed actions, neural computation reflected in
DMN dynamics could be constantly shaped by prediction error through feedback loops.
This computational account of DMN will be described in the mathematical framework of Markov Decision Processes (MDPs). MDPs indeed address specifically the case of decision making in stochastic context with rewards.

Such an RL account of DMN function can naturally embed human behavior
into the tension between exploitative action with immediate gains and
exploratory action with longer-term gratification.

@@

Additionally,
this network was proposed to be situated
at the top of the brain network hierarchy, with
the subordinate salience and dorsal attention network in the middle and
the unimodal sensory cortices at the bottom
\citep{carhart2010default, margulies2016situating}.
Its putative involvement in thinking about the past,
hypothetical experiences, and the future
appears to tie in with the implicit computation of
action and state cascades as a function of what happened in the past.
A policy matrix encapsulates the repertoire of possible actions
on the world given a current state.
The policy matrix encodes the probabilites of
choosing actions to be executed in a certain situation.
The DMN may subserve
  constant exploration of possible action trajectories and
  nested estimaton of their
  cumulative reward outcomes. Implicit computation of future choices
  provides a potential explanation for the
  evolutionary emergence and practical usefulness of
  mind-wandering at day-time and dreams during sleep
  in humans.
\begin{figure}[!h]
  \includegraphics[width=.9\linewidth]{neurobiological_and_rl_overview_DMN.pdf}
  \caption{\textbf{Default mode network:
  neurobiological implementation of reinforcement learning.}
  Overview of how the constituent nodes of the DMN (refer to section \ref{sec:nodes}) may
  map onto computational components necessary for an RL agent.}
  \label{fig:rl_process_chart}
\end{figure}
